---
title: "Data Description"
author: "Göran Broström"
date: "12 June 2018"
output: html_document
---

```{r setup, include=FALSE,comment=NA}
knitr::opts_chunk$set(echo = TRUE,comment=NA)
options(width = 90)
```

# Introduction

The *U18006* data retrieval is tabulated for all persons of exact age 60 years somwhere between 1 November 1960 and 31 December 2013. The table can be downloaded from [my office computer](http://capa.ddb.umu.se). See **R** code below.

# Data

We download the data frame.

```{r getdata}
con <- url("http://capa.ddb.umu.se/data/tab_all.rds")
se <- readRDS(con)
close(con)
head(se)
```
Showing the first six lines of *se*. For all combinations of *AR, KOMMUN, sex, civst, period, hisclass, and age*, the number of deaths (*event*) and exposure time (*exposure*) are calculated. It results in `r NROW(se)` combinations with data. There are of course many empty cells; they are not shown in the data frame.

One problem is the variable *KOMMUN*: There are `r length(unique(se$KOMMUN))` unique *KOMMUN* codes in the data set, but because of some *kommun reforms* over the time, the question is how we should and can utilize this information.

# Cohort information

```{r agebyar, results='asis'}
knitr::kable(with(se, tapply(event, list(AR, age), sum)))
##with(se, tapply(event, list(AR, age), sum))
```

This table shows the number of deaths by age and cohort (year at age 60). *No-data combinations* are represented by *NA*.

Obviously (and as expected) there is only one cohort that can be followed to the end, those aged 60 between 1960 and 1965. A naive analysis for men in this cohort could be like

```{r naivecohort}
fit <- glm(event ~ offset(log(exposure)) + hisclass + age,
           data = se[se$AR == 1960 & se$sex == "male", ], 
           family = poisson())
drop1(fit, test = "Chisq")
round(summary(fit)$coefficients[1:5, c(1, 2)], 3)
```
We see that both *hisclass* and age (surprise?) are highly *statistically* significant. The *none* hisclass stands out with highest mortality (The *elite* is reference). 

# Period information

The analogous period information.

```{r agebyperiod}
##knitr::kable(with(se, tapply(event, list(AR, age), sum)))
with(se, tapply(event, list(period, age), sum))
```

A full period analysis is only available for *2010-2015*. For men in this period we get

```{r naiveperiod}
fit <- glm(event ~ offset(log(exposure)) + hisclass + age,
           data = se[se$period == "2010-2015" & se$sex == "male", ], 
           family = poisson())
drop1(fit, test = "Chisq")
round(summary(fit)$coefficients[1:5, c(1, 2)], 3)
```

Here the *elite* stands out with clearly lowest mortality of all.

# Conclusion

This is a *very preliminary* tabulation of the data in *U18006*. 

*   Do data look reasonable?
*   Future structure of our analyses?